\section{Discussion}

\subsection{Performance Analysis}

Overall, the proposed receipt OCR pipeline demonstrates strong performance across both text
detection and recognition stages. The detection model achieves a precision of 93.83\% and an
F1-score of 91.62\%, indicating that the system reliably localizes text regions in scanned
receipt images with minimal false positives. The IoU distribution further confirms this, with
the majority of predictions falling in the 0.75--0.90 range, reflecting tight and accurate
bounding box alignment with ground truth annotations.

At the recognition stage, the model achieves a character-level accuracy of 87.96\% (CER of
12.04\%), which is a reasonable result given the inherent challenges of receipt imagery.
However, the relatively high WER of 64.89\% and line accuracy of 63.99\% reveal that while
individual characters are often correctly recognized, complete word and line-level transcription
remains difficult. This gap between character-level and word-level accuracy is expected in OCR
systems: a single character error within a word is sufficient to mark the entire word as
incorrect, making WER a considerably stricter metric than CER.

The preprocessing pipeline contributes meaningfully to these results by addressing common
receipt image degradations prior to detection and recognition. Techniques such as CLAHE for
contrast enhancement, bilateral filtering for noise reduction, and adaptive thresholding for
binarization collectively improve the clarity of text regions, particularly in thermal receipts
with faded ink or photos taken under uneven lighting. The blue-ink detection mechanism also
adds robustness by adjusting the grayscale conversion for invoices that use colored ink, which
standard conversion would otherwise handle poorly.

The most frequent character-level errors such as \texttt{O → 0}, \texttt{S → 5}, and
\texttt{0 → 1} are consistent with patterns commonly observed in receipt OCR literature.
These confusions arise because receipts frequently mix alphanumeric characters in close
proximity (e.g., product codes, prices, and invoice numbers), and the small font sizes typical
of thermal printing reduce the visual distinction between similar-looking glyphs.

\subsection{Limitations}

Despite the encouraging results, several limitations are worth acknowledging. First, the
recognition performance degrades noticeably on receipts with low print quality, heavy noise, or
significant skew that preprocessing alone cannot fully correct. While the preprocessing pipeline
handles many common degradation types, extreme cases such as severely creased receipts or
heavily faded thermal paper remain challenging.

Second, the system's recognition accuracy is sensitive to font size and character density.
Receipts that contain small-font footnotes, tightly packed item descriptions, or multi-column
layouts can cause the detection model to either merge adjacent text regions into a single
bounding box or split a single word across multiple boxes, both of which propagate errors into
the recognition stage.

Third, although the SROIE dataset is a well-established benchmark, it consists predominantly
of Malaysian retail receipts. This limits the generalizability of the trained model to other
receipt formats, languages, or regions. Receipts from different countries often follow
different structural conventions, use different character sets, or include non-Latin scripts
that the current model was not trained to handle.

Finally, the current pipeline treats detection and recognition as two independent sequential
stages. Errors made during detection such as a misaligned or missing bounding box, they cannot
be recovered at the recognition stage, meaning early-stage failures have a compounding effect
on the final output quality.

\subsection{Future Improvements}

Several directions could meaningfully improve the system in future work. On the preprocessing
side, incorporating learning-based image restoration models such as deblurring or
super-resolution networks could better handle low-quality inputs that rule-based preprocessing
struggles with. Automatic skew correction using Hough transform-based deskewing or a learned
document rectification network would also reduce a common source of detection errors.

For text detection, experimenting with alternative architectures beyond DBNet++ and EAST
such as PAN++ or TextBPN could improve recall on densely packed or irregularly shaped
text regions. Incorporating end-to-end trainable detection-recognition frameworks, where
both stages are jointly optimized, would eliminate the error propagation bottleneck inherent
in the current two-stage design.

On the recognition side, replacing or augmenting the CTC decoder with an attention-based
sequence decoder could improve performance on ambiguous characters by leveraging broader
contextual information. Additionally, fine-tuning the recognition model on a more diverse
receipt dataset that spans multiple languages, countries, and printing technologies would
significantly improve generalizability.

Finally, integrating a post-processing step such as a language model or dictionary-based
correction layer could resolve common character confusions like \texttt{O/0} and
\texttt{S/5} by leveraging semantic context, for example knowing that a price field should
contain digits rather than letters.