\section{Introduction}

\subsection{Background}
In today's increasingly digitized economy, businesses and individuals generate enormous volumes of paper receipts on a daily basis. From retail stores and restaurants to hospitals and government agencies, receipts serve as critical records of financial transactions. Yet despite the digital revolution transforming nearly every other aspect of commerce, the processing of these physical documents remains surprisingly manual in many organizations. Accountants still key in figures by hand, expense management teams sort through stacks of paper, and auditors cross-reference printed records against digital ledgers — a workflow that is both tedious and prone to human error.
Optical Character Recognition (OCR) has long promised to bridge the gap between the physical and digital worlds by converting printed or handwritten text in images into machine-readable data. However, applying OCR to real-world receipts is far from trivial. Unlike clean, typeset documents, scanned receipts present a host of visual challenges: varying fonts and layouts, low print quality, skewed scanning angles, background noise, and inconsistent formatting across different vendors and regions. Simply recognizing the characters on a receipt is only half the battle — the other half involves understanding the semantic structure of the document well enough to extract meaningful fields such as the store name, transaction date, individual line items, and total amount.
This is where the intersection of Computer Vision and Natural Language Processing becomes particularly powerful. Modern deep learning approaches have dramatically improved both the accuracy of text detection and recognition in natural scene images, and the ability to extract structured information from unstructured document layouts. The ICDAR 2019 Robust Reading Challenge on Scanned Receipts OCR and Information Extraction (SROIE) competition was organized precisely to advance the state of the art in this domain, providing a standardized benchmark dataset and evaluation framework that has since attracted significant research attention.
\subsection{Problem Statement}
Automating the extraction of key information from scanned receipts involves solving two tightly coupled sub-problems. The first is OCR — accurately localizing and transcribing all text regions within a receipt image despite the visual noise and distortions inherent in scanned documents. The second is information extraction — parsing the recognized text to identify and classify specific semantic fields that carry business-relevant meaning, such as the merchant name, date and time of the transaction, and the total amount paid.
Both tasks are non-trivial in isolation, and their combination introduces additional complexity. OCR errors propagate into the information extraction stage, meaning that even a small misrecognition — for instance, confusing the digit "0" with the letter "O," or misreading a price due to ink bleeding — can result in incorrect or missing extracted values. Furthermore, receipts do not follow a universal template. A receipt from a Malaysian bookstore looks structurally very different from one issued by a Starbucks outlet in the United States, yet both must be processed correctly by a robust system. The diversity of receipt layouts, languages, and printing qualities makes it difficult to rely on rigid rule-based parsing, motivating the use of data-driven Computer Vision and machine learning approaches.
\subsection{Objectives}
The primary objectives of this project are as follows:

- To develop and implement a pipeline for receipt OCR and key information extraction using the ICDAR 2019 SROIE dataset.

- To perform text detection and recognition on scanned receipt images, accurately localizing bounding boxes around text regions and transcribing their content.

- To evaluate the performance of the implemented approach using the metrics defined by the SROIE challenge, including precision, recall, and F1-score for both the OCR and information extraction tasks.

- To analyze the challenges encountered in processing real-world receipt images and discuss the limitations and potential improvements of the proposed system.
\subsection{Dataset Description}
This project uses the ICDAR 2019 Scanned Receipts OCR and Information Extraction (SROIE) dataset, made publicly available through the official competition repository and mirrored on GitHub by zzzDavid. The dataset consists of 1,000 whole scanned receipt images. Each image is paired with bounding box annotations for text localization and a JSON file containing four key ground truth fields: company name, date, address, and total amount. The receipts span a wide variety of merchants and layouts, presenting realistic challenges such as skew, noise, and inconsistent formatting that make it a rigorous benchmark for document understanding systems.
\newpage